% Avoiding tipping points in the management of ecological systems: a non-parametric Bayesian approach

<!-- Run R code for analysis, to be called by figures -->
```{r generics-and-libraries, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
library(nonparametricbayes) 
library(ggplot2)
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
opts_chunk$set(external=TRUE)
read_chunk("../examples/BUGS/external-chunks.R")
```

```{r plot-options, cache=FALSE}
options(xtable.print.comment=FALSE)
options(xtable.type = 'latex', table.placement="H")
opts_knit$set(progress = TRUE, verbose = TRUE)
opts_chunk$set(dev = c('pdf', 'png'), fig.width=5.5, fig.height=4,
 cache.path = 'cache/nonparametric-bayes/', cache=TRUE, external=TRUE)
opts_chunk$set(warning=FALSE, message=FALSE, comment=NA,
               tidy=FALSE)
toggle = "markup" # "hide" 
theme_set(theme_bw(base_size=12))
```

```{r posterior-mode}
```
```{r stateeq}
```
```{r sdp-pars, dependson="stateeq"}
```
```{r obs, dependson="sdp-pars"}
```


```{r mle, dependson="obs"}
```
```{r mle-output, dependson="mle", results=toggle}
```


```{r gp-priors}
```
```{r gp, dependson=c("gp-priors", "obs")}
```
```{r gp_traces_densities, dependson="gp"}
```
```{r gp-output, dependson="gp", results=toggle}
```


```{r jags-setup}
```
```{r common-priors}
```
```{r allen-model}
```
```{r allen-priors, dependson="common-priors"}
```
```{r allen-mcmc, dependson=c("allen-model", "allen-pars", "jags-setup"), results=toggle}
```
```{r allen-traces, dependson="allen-mcmc"}
```
```{r allen-posteriors, dependson=c("allen-traces", "allen-priors")}
```
```{r allen-output, dependson=c("posterior-mode", "allen-traces"), results=toggle}
```


```{r ricker-model}
```
```{r ricker-priors, dependson="common-priors"}
```
```{r ricker-mcmc, dependson="ricker-model", results=toggle}
```
```{r ricker-traces, dependson="ricker-mcmc"}
```
```{r ricker-posteriors, dependson=c("ricker-traces", "ricker-priors")}
```
```{r ricker-output, dependson=c("posterior-mode", "ricker-traces"), results=toggle}
```


```{r myers-model}
```
```{r myers-priors}
```
```{r myers-mcmc, dependson="myers-model", results=toggle}
```
```{r myers-traces, dependson="myers-mcmc"}
```
```{r myers-posteriors, dependson="myers-traces"}
```
```{r myers-output, dependson=c("posterior-mode", "myers-traces"), results=toggle}
```


```{r assemble-models, dependson=c("myers-output", "ricker-output", "allen-output", "gp-output", "mle-output")}
```
```{r gp-opt, dependson="gp-output"}
```
```{r mle-opt, dependson="mle-output"}
```
```{r allen-opt, dependson="allen-output"}
```
```{r ricker-opt, dependson="ricker-output"}
```
```{r myers-opt, dependson="myers-output"}
```
```{r assemble-opt, dependson=c("gp-opt", "mle-opt", "allen-opt", "ricker-opt", "myers-opt")}
```
```{r sims, dependson="assemble-opt"}
```


```{r profits, dependson="sims", results=toggle}
```
```{r deviances, dependson=c("posterior-mode", "myers-output", "ricker-output", "allen-output", "gp-output", "mle-output"), results=toggle}
```




Abstract
=======================================================================

Decision-theoretic methods often rely on simple parametric models of
ecological dynamics to compare the value of a potential sequence of
actions. Unfortunately, such simple models rarely capture the complexity
or uncertainty found in most real ecosystems. 

We demonstrate how nonparametric Bayesian models can provide robust,
nearly-optimal solutions to decision making under uncertainty when
_we don't know the correct model_ to use. 


While methods that account
for _parametric_ uncertainty can be very successful with the right model,
structural uncertainty of not knowing what model best approximates the 
dynamics poses considerably greater difficulty.  


Non-parametric Bayesian
methods offer a promising statistical approach for predictive modeling
of ecological dynamics in regions of state space where the data is
adequate, while at the same time offering more flexible patterns with
greater uncertainty outside the observed data. This contrasts from simple
parametric models which provide relatively constant level of uncertainty
in regions with and without adequate data. The consequence of such
misplaced confidence outside the data can lead to highly undesirable
results that may be avoided with the more flexible non-parametric
Bayesian approach.


Introduction
=======================================================================

Decision making under uncertainty is a ubiquitous challenge of natural 
resource management and conservation.  

Here we illustrate how a stochastic dynamic programming algorithm can 
be driven by the predictions from a Gaussian process model, sidestepping
the need for an accurate model-based description of the sustem dynamics.  

<!-- Models for decision-making -->
Decision-theoretic (or optimal control) tools require a model that can assign probabilities of
future states (e.g. stock size of a fishery) given the current state and a
proposed action (e.g. fishing harvest or effort).  The decision maker then
seeks to determining the course of actions (also referred to as the policy)
that maximizes the expected value of some objective function (such as
net present value derived from the resource over time.  (The approach
can be adapted to permit alternatives to maximizing expectation, such
as minimizing the maximum cost or damage that might be incurred. See
@Polasky2011).
Management frequently faces a sequential decision-making problem -- after 
selecting an action, the decision-maker may receive new information about 
the current state and must again choose an appropriate action -- i.e. setting
the harvest limits each year based on stock assessments the year prior.  

<!-- Parametric models -->

Traditional approaches to optimal control (Pontryagin's principle, stochastic
dynamic programming) rely on knowledge of the state equation.  Such simple 
parametric models, frequently justified by mechanistic underpinnings are used
to provide probabilities of future states.  Fisheries literature, which enjoys
some of the longest history in the use of the optimal control framework, provides
a clear example. Models such as the Ricker or Beverton Holt curves are ubiquitous [too strong? references?]
Parametric models can reflect variability introduced by a stochastic environment
or demographic process my incorporating random shocks from a given distribution. 


<!-- Parametric uncertainty -->
As the parameter values for these models must be estimated from limited data,  
there will always be some uncertainty associated with these values.  This uncertainty
further compounds the intrinsic variability introduced by demographic or environmental 
noise.  The degree of uncertainty in the parameter values can be inferred from the data
and reflected in the estimates of the transition probabilities.  

<!-- Structural Uncertainty --> 
Unfortunately, estimates of parameter uncertainty are only as good as the parametric 
models themselves.  Too often, we do not understand the system dynamics well enough
to know if a model provides a good approximation over the relevant range of states
and timescales (criteria that we loosely refer to as defining the "right" or "true" model.)
So called structural or model uncertainty is a more difficult problem than parametric 
uncertainty.  Typical solutions involve either model choice, or model averaging. Either
approach remains limited by the degree to which any of the proposed models are sufficiently
close to the right model.  

<!-- Comparing models -->
The nature of decision-making problems provides a convenient way to compare 
models.  Rather than compare models in terms of best fit to data or fret over
the appropriate penalty for model complexity, model performance is defined 
in the concrete terms of the decision-maker's objective function, which we
will take as given. (Much argument can be made over the 'correct' objective
function, e.g. how to account for the social value of fish left in the sea
vs. the commercial value of fish harvested; see @Halpern2013 for further 
discussion of this issue.  Alternatively, we can always compare model performance
across multiple potential objective functions.)  The decision-maker
does not necessarily need a model that provides the best mechanistic understanding
or the best long-term outcome, but rather the one that best estimates the 
probabilities of being in different states as a result of the possible actions. 



<!-- The weakness of parametric models -->


While simple mechanistic models can nevertheless provide important insights
into long-term outcomes, such approaches are not well-suited for use in
forecasting outcomes of potential management options.  Non-parametric
approaches offer a more flexible alternative that can both more accurately
reflect the data available while also representing greater uncertainty
in areas (of state-space) where data is lacking.

We demonstrate how
a Gaussian Process model of stock recruitment can lead to nearly optimal
management through stochastic dynamic programming, comperable to knowing
the correct structural equation for the underlying simulation.  Meanwhile,
parametric models that do not match the underlying dynamics can perform 
very poorly, even though they fit the data as well as the true model.  

Ecological research and management strategy should pay closer attention
to the opportunities and challenges nonparametric modeling can offer.



<!--
### Quantitative vs Qualitative Decisions

In this paper, we consider those ecological management problems in which
a mathematical (or computational) model is used to quantitatively inform
decision-making by allowing a manager to compare to the expected consequences
of potential management actions (or policies).  We distinguish this from the
solely qualitative use of a model, in which models are used to represent and
compare hypotheses of different mechanisms that would lead to qualitatively 
different actions. 

Though simple mechanistically motivated models
may be best for the latter case [@Cuddington2013; @Geritz2012], such models 
can be not only inaccurate but misleading in quantitative decision making. This 
arises for two reasons:

1. We do not know what the correct parameter values are for the models.
1. We do not know what the correct models are for ecological systems.

The complexity of ecological interactions and a lack of data contribute
greatly to both of the problems. This concern is particularly acute
in case of ecological tipping points [@Barnosky2012; @Scheffer2001],
which arise from feedbacks common in ecological systems and can lead to a
sudden catastrophic transition to an undesirable state.  In general we do
not know where such tipping points are unless we have already observed
the transition, in which case it is frequently too late to respond.
Though there may exist early warning signals for certain kinds of these
transitions that are driven by slow changes [@Scheffer2009], we do not
know when, where, or how to apply them to the decision making context
more generally [@Boettiger2013]. 

In addition to facing the potentially grave consequences of such transitions,
if they do exist in a system of interest, this generally means that we lack
data in the region near and beyond a possible tipping point.  As a consequence,
our models are least accurate just where we need them to be most accurate.  
Unfortunately, parametric modeling approaches are not well-suited to this 
scenario, in which good data is avialble over only part of the relevant 
state-space.  Necessary assumptions about model structure in any parametric
approach can lead such models to perform very poorly, even when accounting
for parameter uncertainty. 

-->

Non-parametric models can better represent uncertainties outside 
of observed data while also better capturing the dynamics in the region
observed.  Advances in the theory and computational implementations of 
nonparametric methods make them ripe for such applications.  We use the classic
problem of optimal harvest of a marine fishery to illustrate how the 
nonparametric approach of Gaussian processes can be applied.  We will 
compare Bayesian implementations of both nonparametric and parametric 
models, which best allow us to capture the uncertainty of model estimation
in either case and permits a more natural comparison between approaches.  


### Terminology: "Non-parametric" 

The term non-parametric to describe modeling approaches such as Gaussian
processes is a common source of confusion, notwithstanding the fact that
the model is still specified by parameters.  Some literature has introduced
the term "semi-parametric" in tacit acknowledgment of this, which no doubt
only contributes to the confusion.  The problem is further exacerbated by
the several meanings assigned to the term in statistics: (a) that the method 
does not assume some particular probability distribution (e.g. a non-parametric bootstrap)
or (b) that the method does not assume a fixed structure to the model.  Our 
use is that of (b), in contrast to the classical approaches that do.  The 
difference is clearest by way of example. 




Approach and Methods
=====================================================================

## Underlying Model

Concerns over the potential for tipping points in ecological dynamics
[@Scheffer2001] highlight the dangers of uncertainty in ecological
management and pose a substantial challenge to existing decision-theoretic
approaches [@Brozovic2011].  Because intervention is often too late 
after a tipping point has been crossed (but see @Hughes2013), management
is most often concerned with avoiding potentially catastrophic tipping
points before any data is available at or following a transition that
would more clearly reveal these regime shift dynamics [e.g. @Bestelmeyer2012].

To illustrate the value of the non-parametric Bayesian approach to management,
we focus on example of a system containing such a tipping point whose dynamics
can still be described by a simple, one-dimensional parametric model.  
We will focus on a simple parametric model for a single species [derived
from fist principles by @Allen2005a] as our underlying "reality".
X_{t+1} &= Z_t f(S_t) \\
S_t &= X_t - h_t \\
f(S_t) &= S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - C\right)}
\end{align}

Where $Z_t$ is multiplicative noise function with mean 1, representing
stochastic growth. We will consider log-normal noise with shape parameter
$\sigma_g$.  We start with an example in which the parameters are $r =2$,
$K=8$, $C=5$ and $\sigma_g = 0.1$.
<!-- ` r p[1]`, $K =$ ` r p[2]`, $C =$ ` r p[3]`, and  $\sigma_g =$ ` r sigma_g`. -->


As a low-dimensional system completely described by three parameters, this
scenario should if anything be favorable to a parametric-based approach.
This model contains an Allee effect, or tipping point, below which the
population is not self-sustaining and shrinks to zero [@Courchamp2008].


Both parametric and nonparametric approaches will require training
data on which to base their model of the process.  We generate the
training data under the model described in Eq 1 for $T_{\textrm{obs}}=40$ time
steps, under a known but not necessarily optimal sequence of harvest
intensities, $h_t$.  For simplicity we imagine a fishery that started
from zero harvest pressure and has been gradually increasing the harvest.


<!-- things that we could hit upon in discussion:
Using data simulated from a specified model rather than empirical data
permits the comparison against the true underlying dynamics, setting 
a bar for the optimal performance possible.  

(Motivation, alternatives, stationarity, examples without a stable node
(limit-cycle models), examples based on observations near a stable node
alone, and why that isn't impossible).
-->

<span style="color:#007FFF;">***Should we include any emprical examples?***</span>

### The optimal control problem in fisheries management

In our example, we focus on the problem in which a manager must set 
the harvest level for a marine fishery each year to maximize the net
present value of the resource, given an estimated stock size from the year
before. The economic value and ecological concern have made marine fisheries the crucible for much
of the founding work [@Gordon1954; @Reed1979; @May1979; @Ludwig1982]
in managing ecosystems under uncertainty.  Global trends [@Worm2006]
and controversy [@Hilborn2007; @Worm2009] have made understanding these
challenges all the more pressing. 

To permit comparisons against a theoretical optimum we will consider
data on the stock dynamics simulated from a simple parametric model
in which recruitment of the fish stock $X_{t+1}$ in the following year
is a stochastic process governed by a function $f$ of the current 
stock $X_t$, selected harvest policy $h_t$, and noise process $Z$,

$$X_{t+1} = Z_t f(X_t, h_t) $$


Given parameters for the function $f$ and probability distribution $Z$,
along with a given economic model determining the  price/profit $\Pi(X_t,
h_t)$ realized in a given year given a choice of harvest $h_t$ and
observed stock $X_t$.  This problem can be solved exactly for discretized
values of stock $X$ and policy $h$ using stochastic dynamic programming
(SDP) [@Mangel1985]. Problems of this sort underpin much marine fisheries
management today.

A crux of this approach is correctly specifying the functional form of $f$,
along with its parameters.  The standard approach uses one of a 
handful of common parametric models representing the stock-recruitment
relationship, usually after estimating the model parameters from any 
available existing data. Uncertainty in the parameter estimates can 
be estimated and integrated over to determine the optimal policy under
under uncertainty [@Mangel1985; @Schapaugh2013]. Uncertainty in the model
structure itself can only be addressed in this approach by hypothesizing 
alternative model structures, and then performing some model choice or
model averaging  [@Williams2001; @Athanassoglou2012]. 


### Parametric Models

We consider three candidate parametric models for the stock-recruitment function, 
which we refer to by the first authors of the publications in which they were
first proposed.  

We generate the data with a four-parameter model that contains
a tipping point (an Allee effect, see [@Allen, @Courchamp]) below which the stock
decreases to zero, 

$$X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K}\right)\left(\frac{S_t- \theta}{K}\right)} \label{Allen} $$

$$ S_t = X_t - h_t $$

The parameter $C$ reflects the location of the tipping point, $K$ the carrying capacity of
the stock, and $r$ the base recruitment rate.  $S_t$ represents the stock size after
a harvest $h_t$ has been implemented.  $Z_t$ represents a log-normal random variable of
log-mean zero and log-standard deviation parameter $\sigma$.  


We consider two alternative candidate models: the Ricker [@Ricker] stock-recruitment curve,

$$X_{t+1} = Z_t X_t e^{r \left(1 - \frac{S_t}{K} \right) } \label{Ricker} $$

and an alternative four-parameter model adapted from @Myers, 

$$ X_{t+1} = Z_t \frac{r S_t^{\theta}}{1 - \frac{S_t^\theta}{K}} \label{Myers} $$

which contains a tipping point for $\theta > 2$ and becomes a Beverton-Holt model 
at $\theta = 1$.  

<!-- Note that while similarly-named parameters contain similar connotations, the 
estimate and precise intepretation will be different in each model. -->


### Bayesian Inference of Parametric models 

Given the sample data, we infer posterior distributions for each of the three
models listed above using a Markov Chain Monte Carlo Gibbs Sampler (jags, see 
appendix for implementation details and code) given uniform priors.  We run six
chains for $10^6$ steps each and then assess convergence by Gelman-Rubin criterion
and inspection of the traces, see appendix.  

<!-- By transforming variables we can integrate out the the $\sigma$ parameter
and the constant term in the log-mean, see appendix.  The resulting one or
two parameters can then be evaluated over a vector or grid of possible values,
and the result transformed back to determine the posterior density of the parameters,
This approach avoids any concerns of convergence in the MCMC analysis and can thus
be used to verify that the performance of the parametric models can not be explained
by convergence issues.  However, this approach does not generalize easily to richer models,
such as the introduction of measurement noise or additional parameters, in which 
MCMC based approaches are the obvious choice.  -->




### The Non-parametric Bayesian alternative for stock-recruitment curves

The use of Gaussian process (GP) regression (or "kriging" in the geospatial
literature) to formulate a predictive model is relatively new in the
context of modeling dynamical systems [@Kocijan2005] and introduced
in the ecological modeling and fisheries management by @Munch2005.
An accessible and thorough introduction to the formulation and use of
GPs can be found in @Rasmussen2006.

The essence of the GP approach can be captured in the
following thought experiment: An exhaustive parametric approach to the
challenge of structural uncertainty might proceed by writing down all
possible functional forms for the underlying dynamical system with all
possible parameter values for each form, and then consider searching
over this huge space to select the most likely model and parameters;
or using a Bayesian approach, assign priors to each of these possible
models and infer the posterior distribution of possible models. The
GP approach can be thought of as a computationally
efficient approximation to this approach. GPs represent
a large class of models that can be though of as capturing or reasonably
approximating the set of models in this collection.  By modeling at the
level of the process, rather than the level of parametric equation,
we can more concisely capture the possible behavior of these curves.
In place of a parametric model of the dynamical system, the GP 
approach postulates a prior distribution of (n-dimensional)
curves that can be though of as approximations to a range of possible
(parametric) models that might describe the data. The GP allows us
to consider a set of possible curves simultaneously.  



### Background on Gaussian Process inference

Once the posterior Gaussian process (GP) has been estimated [e.g. see
@Munch2005], it is necessary to adapt it in place of the parametric
equation for the stochastic dynamic programming (SDP) solution [see
@Mangel1988 for a detailed description of parametric SDP methods] to the
optimal policy. The essence of the idea is straight forward -- we will use
the estimated GP in place of the parametric growth function to determine
the stochastic transition matrix on which the SDP calculations are based.
The SDP is solved in a discretized state space -- both the continuously 
valued population densities $X$ and harvest quotas $h$ are first mapped to
a bounded, discrete grid.  (For simplicity we will consider a uniform grid,
though for either parametric or GP-based SDP it is often advantageous to 
use a non-uniform discretization such as a basis function representation,
e.g. see [@Deisenroth2009]). 

The SDP approach then computes a transition matrix, $\mathbf{F}$.  We 
demonstrate that calculation is just as straight forward based on the GP
as it is in the classical context using the parametric model.  The 
${i,j}$ of the transition matrix $F$ entry gives the probability of transitioning into 
state $x_i$ given that the system is in state $x_j$ in the previous 
time-step.  To generate the transition
matrix based on the posterior GP, we need only the expected values
at each grid point and the corresponding variances (the diagonal of
the covariance matrix), as shown in Figure 1.  Given the mean of the 
GP posterior at each grid-point as the vector $E$ and variance at that
point as vector $V$, the probability of transitioning from state $x_i$ to state $x_j$ is

$$\mathcal{N}\left(x_j | \mu = E_i, \sigma = \sqrt{V_i}\right)$$

where $\mathcal{N}$ is the Normal density at $x_j$ with mean $\mu$ and
variance $\sigma^2$.  Strictly speaking, the transition probability should
be calculated by integrating the normal density over the bin of width
$\Delta$ centered at $x_j$.  For a sufficiently fine grid that $f(x_j)
\approx f(x_j + \Delta)$, it is sufficient to calculate the density at
$x_j$ and then row-normalize the transition matrix. The process
can then be repeated for each possible discrete value of our control 
variable, (harvest $h$).  


**Pseudocode for the determining the transition matrix from the GP**

```r
for(h in h_grid)
  F_h = for(x_j in grid)
          for(i in 1:N) 
            dnorm(x_j, mu[i]-h, V[i])
```


Using the discrete transition matrix we may write down the Bellman
recursion defining the stochastic dynamic programming iteration:

\begin{equation}
V_t(x_t) = \max_h \mathbf{E} \left( h_t + \delta V_{t+1}( Z_{t+1} f(x_t - h_t)) \right)
\end{equation}

where $V(x_t)$ is the value of being at state $x$ at time $t$, $h$
is control (harvest level) chosen. Numerically, the maximization is
accomplished as follows. Consider the set of possible control values to
be the discrete values corresponding the grid of
stock sizes.  Then for each $h_t$ there is a corresponding transition
matrix $\mathbf{F}_h$ determined as described above but with mean 
$\mu = x_j - h_t$. Let $\vec{V_t}$ be the vector whose $i$th element corresponds
to the value of having stock $x_i$ at time $t$.  Then let $\Pi_h$ be
the vector whose $i$th element indicates the profit from harvesting
at intensity $h_t$ given a population $x_i$ (e.g. $\max(x_i, h_t)$
since one cannot harvest more fish then the current population size).
Then the Bellman recursion can be given in matrix form as

$$V_{t} = \max_h \left( \Pi_{h_{t}} + \delta \mathbf{F}_h V_{t+1} \right)$$

where the sum is element by element and the expectation is computed by the matrix multiplication $\mathbf{F} V_{t+1}$.  

### Pseudocode for the Bellman iteration

```r
 V1 <- sapply(1:length(h_grid), function(h){
      delta * F[[h]] %*% V +  profit(x_grid, h_grid[h]) 
    })
    # find havest, h that gives the maximum value
    out <- sapply(1:gridsize, function(j){
      value <- max(V1[j,], na.rm = T) # each col is a diff h, max over these
      index <- which.max(V1[j,])  # store index so we can recover h's 
      c(value, index) # returns both profit value & index of optimal h.  
    })
    # Sets V[t+1] = max_h V[t] at each possible state value, x
    V <- out[1,]                        # The new value-to-go
    D[,OptTime-time+1] <- out[2,]       # The index positions
```

This completes the algorithm adapting the GP to the sequential decision-making 
problem through SDP, which we believe has not yet been demonstrated in the
literature.  We provide an R package implementation of this, along with
the Gaussian process inference, in the supplemental materials.  


### Estimating parametric models 

We estimate posterior distributions for two parametric models: one using the structurally correct
model as given in Eq (1), which we refer to as the "Parametric Bayes" model,
and another using the familiar Ricker model, using a Gibbs sampler
as described (with source code) in the appendix).  In addition we 
estimate the parameters of the structurally correct model by maximum
likelihood.  

The posterior distribution for the hyper-parameters of the Gaussian 
process model are estimated by Metropolis-Hastings algorithm, again with
details and code provided in the Appendix.  @Rasmussen2006 provides
an excellent general introduction to Gaussian Processes and @Munch2005 
first discusses their application in this context.  

Results
====================================================


```{r Figure1, dependson=c("assemble-models", "plot-options"), fig.cap="Graph of the inferred Gaussian process compared to the true process and maximum-likelihood estimated process.  Graph shows the expected value for the function $f$ under each model.  Two standard deviations from the estimated Gaussian process covariance with (light grey) and without (darker grey) measurement error are also shown.  The training data is also shown as black points.  (The GP is conditioned on 0,0, shown as a pseudo-data point). ", include=FALSE}
```
```{r Figureb, dependson=c("assemble-models", "par-fns", "plot-options"), toggle = 1}
```



Figure 1 shows the mean inferred state space dynamics of each model
relative to the true model used to generate the data, predicting the
relationship between observed stock size (x-axis) to the stock size
after recruitment the following year.  All models except the MLE model
estimate a distribution around the means shown here, and all models 
estimate a level of process noise, which is independent of the state
value (x).  Note that in contrast to the  other models shown, the mean
Gaussian process corresponds to a distribution of curves - as indicated
by the gray band - which itself has a mean shown in black.  Note that 
this mean GP is thus more certain of the dynamics in the region where
data is available then where it is not.  

While it would be natural (and straight forward) to condition
the GP on passing through the origin (0,0) (see appendix), the estimate
shown here is based only on the observed data. The observed data from
which each model is estimated is also shown.  The observations come
from only a limited region of state space corresponding to unharvested
or weakly harvested system.  No observations occur at the theoretical
optimum or near the tipping point.

```{r Figure2, dependson = c("assemble-opt", "plot-options"), fig.cap="The steady-state optimal policy (infinite boundary) calculated under each model.  Policies are shown in terms of target escapement, $S_t$, as under models such as this a constant escapement policy is expected to be optimal [@Reed1979].", toggle = 1}
```

The resulting optimal management strategy based on each of the inferred
models is shown in Figure 2, against the optimal strategy given the
true underlying dynamics.  Policies are shown in terms of target
escapement, $S_t$.  Under models such as this a constant escapement
policy is expected to be optimal [@Reed1979], whereby population levels
below a certain size $S$ are unharvested, while above that size the harvest
strategy aims to return the population to $S$, resulting in the hockey-stick
shaped policies shown.  


```{r Figure3, dependson=c("sim", "plot-options"), fig.cap="Gaussian process inference outperforms parametric estimates. Shown are 100 replicate simulations of the stock dynamics (eq 1) under the policies derived from each of the estimated models, as well as the policy based on the exact underlying model.", toggle = 1}
```

The consequences of managing 100 replicate realizations of the 
simulated fishery under each of the policies estimated is shown in Figure 3.  As expected
from the policy curves, the structurally correct model under-harvests,
leaving the stock to vary around it's un-fished optimum.  The structurally
incorrect Ricker model over-harvests the population passed
the tipping point consistently, resulting in the immediate crash of the stock and 
thus derives minimal profits.  


The results shown in Figures 1-3 are not unique to the simulated data or models chosen
here, but arises across a range of parameter values and simulations as shown in the 
supplemental figures.  The results across this range can most easily be compared 
by the relative differences in net present value realized by each of the approaches,
as shown in Figure 4.  The Gaussian Process most consistently realizes a value 
close to the optimal solution, and importantly avoids ever driving the system across
the tipping point, which results in the near-zero value cases in the parametric models.  


```{r Figure4, dependson=c("profits", "plot-options"), fig.cap="Histograms of the realized net present value of the fishery over a range of simulated data and resulting parameter estimates. For each data set, the three models are estimated as described above. Values plotted are the averages of a given policy over 100 replicate simulations. Details and code provided in the supplement.", toggle = 1}
```

Discussion 
==================================================================


Non-parametric Bayesian methods have received far too little attention
in ecological modeling efforts that are aimed at improved conservation
planning and decision making support.  Such approaches may be particularly
useful when the available data is restricted to a limited area of
state-space, which can lead parametric models to underestimate the
uncertainty in dynamics at population levels (states) which have not
been observed.  One reason for the relative absence of nonparametric
approaches in the natural resource management context may be the lack
of existing approaches for adapting the non-parametric Bayesian models
previously proposed [@Munch2005] to a decision-theoretic framework.
Adapting a non-parametric approach requires modification of existing
methods for decision theory.  We have illustrated how this might be
done for a classic stochastic dynamic programming problem, opening the
door for substantial further research into how these applications might
be improved.


While non-parametric Bayesian approaches will not always be preferable 
to simple mechanistic models, we highlight three aspects of the problem
consider here that make these methods particularly valuable.  These aspects
are common to many conservation decision making problems, which thus merit
greater use of non-parametric approaches that can best take advantage of them.  


### 1. Large uncertainty where the data is poor 

The parametric models perform worst when they propose a management strategy
outside the range of the observed data. The non-parametric Bayesian approach, 
in contrast, allows a predictive model that expresses a great deal of uncertainty
about the probable dynamics outside the observed range, while retaining very
good predictive accuracy in the range observed.  The management policy 
dictated by the GP balance this uncertainty against
the immediate value of the harvest, and act to stabilize the population 
dynamics in a region of state space in which the predictions can be 
reliably reflected by the data.  


### 2. Predictive accuracy where data is good

While expressing larger uncertainty outside the observed data, the GP
can also provide a better fit with smaller uncertainty inside the range
of the observed data. This arises from the greater flexibility of the 
Gaussian process, which describes a large family of possible curves.  
While in a parametric context this over-fitting would be more worrisome --
a high-degree polynomial could fit the data even better -- those 
concerns are driven by the resulting parametric fit outside the data,
which may involve wild oscillations unsupported by the data.  As we 
have seen in #1, the GP is less vulnerable to such unjustified predictions
outside the data, and is meanwhile free to benefit from the greater 
fit where the data is available.  


## Future directions

In this simulated example, the underlying
dynamics are truly governed by a simple parametric model, allowing
the parametric approaches to be more accurate.  Similarly, because the
dynamics are  one-dimensional dynamics and lead to  stable nodes (rather
than other attractors such as limit-cycles resulting in oscillations),
the training data provides relatively limited information about the
dynamics.  For these reasons, we anticipate that in higher-dimensional
examples characteristic of ecosystem management problems that the machine
learning approach will prove even more valuable.

<span style="color:#007FFF;"> Data complexity. Perhaps too far out of scope... </span>
The nonparametric Bayesian approach is also better suited to complex and
disparate data.  Incorporating various sources of information into
mechanistic models can be an immensely difficult due to the increased
complexity involved.  


### Online learning

In our treatment here we have ignored the possibility of learning during the 
management phase, in which the additional observations of the stock size could
potentially improve parameter estimates.  While we intend to address this 
possibility in future work in the context of these non-parametric models,
we have not addressed it here for pedagogical reasons. In the context presented
here, it is clear that the differences in performance arise from differences
in the uncertainty inherent in the model formulations, rather than from 
differing abilities to learn.  Because we consider a threshold system, 
online learning would not change this generic feature of a lack of data in a
certain range of the state space which is better captured by the Gaussian process. 




Acknowledgments
================

This work was partially supported by the Center for Stock Assessment Research, a partnership between the University of California Santa Cruz and the Fisheries Ecology Division, Southwest Fisheries Science Center, Santa Cruz, CA and by NSF grant EF-0924195 to MM.



Appendix
========

The appendices have not yet been assembled.  Meanwhile, code to repeat the analyses, along with a complete log of all research conducted on this project, can be found at: [https://github.com/cboettig/nonparametric-bayes](https://github.com/cboettig/nonparametric-bayes/)


## Markov Chain Monte Carlo Analysis

```{r appendixplots, dependson="assemble-models"}
```


References
==========


