Reed (1979) with Gaussian processes 
========================================================


```{r graphing-options, include=FALSE}
require(pdgControl)
require(nonparametricbayes)
require(ggplot2)
opts_knit$set(upload.fun = socialR::notebook.url)
opts_chunk$set(dev.args=list(bg="transparent"), comment=NA)
theme_set(theme_bw())
theme_update(panel.background = element_rect(fill = "transparent",colour = NA),
             plot.background = element_rect(fill = "transparent",colour = NA))
```


```{r results="hide"}
knit("beverton_holt_data.Rmd")
```

Predict the function over the target grid

```{r}
library(nonparametricbayes)
gp <- gp_fit(obs, X, c(sigma_n=1, l=1.5))
plot.gpfit(gp)
```

Solution using `kernlab`

```{r}
library(kernlab)
#gp <- gausspr(obs$x, obs$y, kernel="rbfdot", kpar=list(sigma=0.5), fit=FALSE, scaled=FALSE, var=.09)
gp <- gausspr(obs$x, obs$y, kernel="rbfdot", scaled=FALSE, var=sigma_g^2)
Ef <- predict(gp, X)
```

Manually get the covariance, $K(x_*, x_*) - K(x_*, x)(K(x, x) - \sigma_n^2\mathbb{I})^{-1}K(x,x_*)$. 

```{r}
var <- gp@kcall$var
var <- 9
Inv <- solve(kernelMatrix(kernelf(gp), xmatrix(gp)) + diag(rep(var, length = length(xmatrix(gp)))))
Cf <- kernelMatrix(kernelf(gp), as.matrix(X)) - 
  kernelMult(kernelf(gp), as.matrix(X), xmatrix(gp),  Inv) %*% 
  kernelMatrix(kernelf(gp), xmatrix(gp), as.matrix(X))
```

```{r}
require(ggplot2)
dat <- data.frame(x=X, y=(Ef), ymin=(Ef-2*sqrt(diag(Cf))), ymax=(Ef+2*sqrt(diag(Cf))))    
ggplot(dat) + 
  geom_ribbon(aes(x=x,y=y, ymin=ymin, ymax=ymax), fill="grey80") +
  geom_line(aes(x=x,y=y), size=1) +
  geom_point(data=obs,aes(x=x,y=y)) + 
  scale_y_continuous(name="output, f(x)") + xlab("input, x") 
```



## Stochastic Dynamic programming solution based on the posterior Gaussian process:

```{r}
rownorm <- function(M) t(apply(M, 1, function(x) x/sum(x)))
h_grid <- x_grid
```

Define a transition matrix $F$ from the Gaussian process, giving the probability of going from state $x_t$ to $x_{t+1}$.
We already have the Gaussian process mean and variance predicted for each point $x$ on our grid, so this is simply:


```{r}
V <- sqrt(diag(Cf))
matrices_gp <- lapply(h_grid, function(h){
  F <- sapply(x_grid, function(x) dnorm(x, Ef-h, V))
  F <- rownorm(F)
})
```

True $f(x)$

```{r}
matrices_true <- lapply(h_grid, function(h){
  mu <- sapply(x_grid, f, h, p)
  F_true <- sapply(x_grid, function(x) dnorm(x, mu, sigma_g))
  F_true <- rownorm(F_true)
})
```


Calculate the policy function using the true F:

```{r}
opt_true <- find_dp_optim(matrices_true, x_grid, h_grid, 20, 0, profit, delta=.01)
```

Calculate using the inferred GP

```{r}
opt_gp <- find_dp_optim(matrices_gp, x_grid, h_grid, 20, 0, profit, delta=.01)
```



```{r policy_plot}
require(reshape2)
policies <- melt(data.frame(stock=x_grid, GP = x_grid[opt_gp$D[,1]], Exact = x_grid[opt_true$D[,1]]), id="stock")
policy_plot <- ggplot(policies, aes(stock, stock - value, color=variable)) + geom_point() + xlab("stock size") + ylab("escapement") 
policy_plot
```

We can see what happens when we attempt to manage a stock using this:

``` {r othernoise}
z_g <- function() rlnorm(1,0, sigma_g)
````

``` {r  simulate }
set.seed(1)
sim_gp <- ForwardSimulate(f, p, x_grid, h_grid, K, opt_gp$D, z_g, profit=profit)
set.seed(1)
sim_true <- ForwardSimulate(f, p, x_grid, h_grid, K, opt_true$D, z_g, profit=profit)
````

``` {r  simplot}
df <- data.frame(time = sim_gp$time, stock_gp = sim_gp$fishstock, stock_true = sim_true$fishstock, harvest_gp = sim_gp$harvest, havest_true = sim_true$harvest)
df <- melt(df, id="time")
simplot <- ggplot(df) + geom_line(aes(time,value, color=variable))
simplot
````

Total Profits

```{r}
sum(sim_gp$profit)
sum(sim_true$profit)
```

